Automatic Matching of Bullet Land Impressions

### 2017-06-28 submission/imaging-paper.tex
Resubmission again!

\linenumbers
%\tableofcontents
%\newpage

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land impression (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land impression. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\label{fig:allx3p}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step1}Step 1 of identifying shoulder locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step1-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step3}Step 2 of identifying shoulder locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$. The orange rectangle shows an example of the smoothing window. Valleys and peaks are detected, if they are not within the same window.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step3-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying shoulder locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside the blue boundaries are removed (shown in grey)}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/loess_step47-1}

@@

\subsection{Horizontal alignment}
Signatures of each of the two land impressions, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are prominent and, again, suggest a match between the land impressions. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land impression signatures.\label{fig:crosscut}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/crosscuts-1}

\end{knitrout}
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/crosscutsX-1}

@@

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.75\textwidth]{figure/smoothfac1-1}

@@

\section{Evaluation}
In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance using the James Hamby study data. As a first step, we automatically assess the quality of each of the land impressions by checking that we can identify a stable region. For this, we compute the cross-correlation of signatures extracted from heights 25$\mu m$ apart. For a stable region, we require a minimum of 0.95 for the cross correlation. Four land impressions from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these land impressions has scratch marks across the surface, also known as `tank rash' \citep{hamby:2009}.
%
\begin{figure}
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Barrel 6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read_x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\textwidth]{images/br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Barrel 9 Bullet 2-4}
\includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unknown Bullet B-2}
\includegraphics[width=\textwidth]{images/b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unknown Bullet Q-4}
\includegraphics[width=\textwidth]{images/q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}Images of the four land impressions that got flagged during the quality assessment. All of them show scratch marks (tank rash) across the striation marks from the barrel. They are excluded from the remainder of the analysis.}
\end{figure}
%
We exclude these four land impressions from further matching considerations and
 run all remaining land impressions from the unknown bullets against all remaining land impressions from known bullets for matches, i.e. we are comparing $15 \times 6 -2 = 90 - 2 = 88$ land impressions from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ land impressions from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM). 
Ideally, results look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land impression C-3 and all 118 land impressions from known bullets. Two land impressions show a high CMS. These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and~\ref{fig:hamby-perfect}c. 
 Unfortunately, not all results are as clear cut. 
It might not be reasonable to assume that we can match all land impressions, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.

\begin{figure}[hbtp]
\begin{subfigure}[t]{\textwidth}\centering
\caption{Maximal number of CMS between unknown bullet land impression C-3 and all of the other 118 considered (known) land impressions. For two land impressions the number of maximum CMS is high. }
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{figure/cms-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land impression with the top matching CMS.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/top-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Top 2 match with C-3 based on CMS.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/top2-1}

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of land impressions with at least thirteen CMS in common are matches. 
%
\begin{figure}[hbtp]
  \centering
%' \begin{subfigure}[b]{\textwidth}\centering
%' \caption{}
\begin{minipage}[t]{.47\textwidth}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/cms-bars-1}

\end{knitrout}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}\centering
\end{minipage}
\begin{minipage}[t]{.52\textwidth}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{figure/cms-spines-1}

@@

Another benefit of the digitized version of the images is that we can apply several hundred decision trees to combine in a random forest~\citep{breiman:2001, randomForest}.  For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third is used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate. Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error. 
Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate for 300 trees. 
%
\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.7\textwidth]{figure/oob-1}

\end{knitrout}
\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
\end{figure}
%
After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039. This is a weighted average between false positive error rate of 0.0001 and an error rate of false negatives of 0.2267. This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
Note that this error rate is based on land-to-land comparisons and is expected to be much lower for bullet-to-bullet comparisons. In the case of the Hamby data, even a single tree results in an overall error rate of zero, if we require that a match of two bullets occurs when at least two of the bullet's land impressions are matched. % (see Figure~\ref{fig:treeresult}). 
%' \begin{figure}[hbtp]
%'   \centering
%' <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
%' bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
%' bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
%'   n = n(),
%'   pred = sum(pred > 0.5),
%'   forest = sum(forest > 0.5)
%' )
%' library(reshape2)
%' bm <- melt(data=bullets, id.var=c("bullet", "n"))
%' 
%' bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
%' qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
%'       shape=prediction, colour=prediction, alpha=I(0.6)) + 
%'   theme_bw()  + ylab("") + 
%' #  scale_shape_manual(values=c(1,2)) +
%'   scale_x_continuous("Percent of correctly predicted land-to-land matches", 
%'                      breaks = c(0,25,50,75,100), limits=c(0,100)) +
%'   theme(legend.position = "bottom")
%' @
%' \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11). Both results for the tree and the random forest are shown. In the tree, at least \hh{40\% of all possible} lands of a bullet are  matched correctly, while the random forest achieves perfect results. Unknown bullet B is the hardest for the tree to match.}
%' \end{figure}
%' %
This makes the errors in the automated approach smaller than the human error in the Hamby study. Out of the 507 participants who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, corresponding to a rate of 0.0011. 
%\hh{XXX calling this a false negative rate is problematic. FEs will be up in arms, because inconclusive doesn't necessarily mean something was missed, it could also mean that there was something wrong with the bullets. XXX }

For the Hamby data, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger experiment. 
Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on a threshold fixed at 0.5 may not be the best approach. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be unnecessary in the Hamby study, because, here, the results from the random forest are very clear cut. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest. As expected, the random forest provides a more realistic estimate of the uncertainty in the classification. %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match. The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.} 
%Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.6\textwidth]{figure/tree:forest-1}

\end{knitrout}
\caption{\label{fig:tree-forest}Prediction results from the tree and the forest. Using a cut-off probability of 0.5 the forest correctly predicts every single comparison. Compared to the tree, the forest's prediction probabilities are  shrunk towards either end of the prediction range.}
\end{figure}

Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to Supplement Section~\ref{supp:randomforest}). 
\begin{figure}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.7\textwidth]{figure/unnamed-chunk-1-1}

>>>>

%\linenumbers
%\tableofcontents
%\newpage

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{side-sketch.png}
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land impression (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land impression. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\label{fig:allx3p}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step1}Step 1 of identifying shoulder locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{loess_step1-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step3}Step 2 of identifying shoulder locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$. The orange rectangle shows an example of the smoothing window. Valleys and peaks are detected, if they are not within the same window.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{loess_step3-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying shoulder locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside the blue boundaries are removed (shown in grey)}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{loess_step47-1}

@@

\subsection{Horizontal alignment}
Signatures of each of the two land impressions, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are prominent and, again, suggest a match between the land impressions. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures.
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land impression signatures.\label{fig:crosscut}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{crosscuts-1}

\end{knitrout}
    }
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{crosscutsX-1}

@@

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.75\textwidth]{smoothfac1-1}

@@

\section{Evaluation}
In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance using the James Hamby study data. As a first step, we automatically assess the quality of each of the land impressions by checking that we can identify a stable region. For this, we compute the cross-correlation of signatures extracted from heights 25$\mu m$ apart. For a stable region, we require a minimum of 0.95 for the cross correlation. Four land impressions from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these land impressions has scratch marks across the surface, also known as `tank rash' \citep{hamby:2009}.
%
\begin{figure}
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Barrel 6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read_x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\textwidth]{br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Barrel 9 Bullet 2-4}
\includegraphics[width=\textwidth]{br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unknown Bullet B-2}
\includegraphics[width=\textwidth]{b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unknown Bullet Q-4}
\includegraphics[width=\textwidth]{q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}Images of the four land impressions that got flagged during the quality assessment. All of them show scratch marks (tank rash) across the striation marks from the barrel. They are excluded from the remainder of the analysis.}
\end{figure}
%
We exclude these four land impressions from further matching considerations and
 run all remaining land impressions from the unknown bullets against all remaining land impressions from known bullets for matches, i.e. we are comparing $15 \times 6 -2 = 90 - 2 = 88$ land impressions from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ land impressions from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
Ideally, results look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land impression C-3 and all 118 land impressions from known bullets. Two land impressions show a high CMS. These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and~\ref{fig:hamby-perfect}c.
 Unfortunately, not all results are as clear cut.
It might not be reasonable to assume that we can match all land impressions, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.

\begin{figure}[hbtp]
\begin{subfigure}[t]{\textwidth}\centering
\caption{Maximal number of CMS between unknown bullet land impression C-3 and all of the other 118 considered (known) land impressions. For two land impressions the number of maximum CMS is high. }
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.5\textwidth]{cms-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land impression with the top matching CMS.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{top-1}

\end{knitrout}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Top 2 match with C-3 based on CMS.}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{top2-1}

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of land impressions with at least thirteen CMS in common are matches.
%
\begin{figure}[hbtp]
  \centering
%' \begin{subfigure}[b]{\textwidth}\centering
%' \caption{}
\begin{minipage}[t]{.47\textwidth}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{cms-bars-1}

\end{knitrout}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}\centering
\end{minipage}
\begin{minipage}[t]{.52\textwidth}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\textwidth]{cms-spines-1}

@@

Another benefit of the digitized version of the images is that we can apply several hundred decision trees to combine in a random forest~\citep{breiman:2001, randomForest}.  For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third is used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate. Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error.
Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate for 300 trees.
%
\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.7\textwidth]{oob-1}

\end{knitrout}
\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
\end{figure}
%
After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039. This is a weighted average between false positive error rate of 0.0001 and an error rate of false negatives of 0.2267. This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:treeforest}).
Note that this error rate is based on land-to-land comparisons and is expected to be much lower for bullet-to-bullet comparisons. In the case of the Hamby data, even a single tree results in an overall error rate of zero, if we require that a match of two bullets occurs when at least two of the bullet's land impressions are matched. % (see Figure~\ref{fig:treeresult}).
%' \begin{figure}[hbtp]
%'   \centering
%' <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
%' bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
%' bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
%'   n = n(),
%'   pred = sum(pred > 0.5),
%'   forest = sum(forest > 0.5)
%' )
%' library(reshape2)
%' bm <- melt(data=bullets, id.var=c("bullet", "n"))
%'
%' bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
%' qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5),
%'       shape=prediction, colour=prediction, alpha=I(0.6)) +
%'   theme_bw()  + ylab("") +
%' #  scale_shape_manual(values=c(1,2)) +
%'   scale_x_continuous("Percent of correctly predicted land-to-land matches",
%'                      breaks = c(0,25,50,75,100), limits=c(0,100)) +
%'   theme(legend.position = "bottom")
%' @
%' \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11). Both results for the tree and the random forest are shown. In the tree, at least \hh{40\% of all possible} lands of a bullet are  matched correctly, while the random forest achieves perfect results. Unknown bullet B is the hardest for the tree to match.}
%' \end{figure}
%' %
This makes the errors in the automated approach smaller than the human error in the Hamby study. Out of the 507 participants who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, corresponding to a rate of 0.0011.
%\hh{XXX calling this a false negative rate is problematic. FEs will be up in arms, because inconclusive doesn't necessarily mean something was missed, it could also mean that there was something wrong with the bullets. XXX }

For the Hamby data, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger experiment.
Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on a threshold fixed at 0.5 may not be the best approach. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be unnecessary in the Hamby study, because, here, the results from the random forest are very clear cut. Figure~\ref{fig:treeforest} shows a comparison of the predicted probabilities of a match by the tree and the random forest. As expected, the random forest provides a more realistic estimate of the uncertainty in the classification. %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match. The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.}
%Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.

\begin{figure}[hbtp]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.6\textwidth]{treeforest-1}

\end{knitrout}
\caption{\label{fig:treeforest}Prediction results from the tree and the forest. Using a cut-off probability of 0.5 the forest correctly predicts every single comparison. Compared to the tree, the forest's prediction probabilities are  shrunk towards either end of the prediction range.}
\end{figure}

Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to Supplement Section~\ref{supp:randomforest}).
\begin{figure}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.7\textwidth]{unnamed-chunk-1-1}
